---
title: "Analysis"
output:
  word_document: default
  html_document: default
---
# Libraries
```{r setup, include=T, warning=F}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stargazer)
library(ggplot2)
library(descr)
library(sjPlot)
library(readxl)
library(stringr)
library(plm)
library(table1)
library(RColorBrewer)
library(lattice)
library(foreign)
library(MASS)
library(Hmisc)
library(reshape2)
options(scipen=999)
```

# Data set
```{r, include=T, warning=F}
load("/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Data/Final Data.Rda")
```

# OLS Regressions

## Na誰ve model
```{r, warning=F}
Naive_OLS_Mat <- lm(data = dfinal, Performance_Math ~ Full_School_Day)
Naive_OLS_Lan <- lm(data = dfinal, Performance_Language ~ Full_School_Day)
Naive_OLS_Pool <- lm(data = dfinal, Performance_MatLang ~ Full_School_Day)

stargazer(Naive_OLS_Mat, Naive_OLS_Lan, Naive_OLS_Pool,
          title = "Table 3. OLS - Effects of full schoool days on test scores (na誰ve estimators)",
          column.labels = c("Mathematics", "Language", "Pooled performance"),
          dep.var.labels = c("Test scores", "Test scores", "Test scores"),
          covariate.labels = "Full-day schooling",
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis). Reference category is Non full-day schooling.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/Naive_OLS.html")
```

## Model with all controls
```{r, warning=F}
Control_OLS_Mat <- lm(data = dfinal, Performance_Math ~ Full_School_Day + Sector_School + School_Area + School_SES + S_Quality + Gender + Student_SES + Edu_Mother + Edu_Father)
Control_OLS_Lan <- lm(data = dfinal, Performance_Language ~ Full_School_Day + Sector_School + School_Area + School_SES + S_Quality + Gender + Student_SES + Edu_Mother + Edu_Father)
Control_OLS_Pool <- lm(data = dfinal, Performance_MatLang ~ Full_School_Day + Sector_School + School_Area + School_SES + S_Quality + Gender + Student_SES + Edu_Mother + Edu_Father)

stargazer(Control_OLS_Mat, Control_OLS_Lan, Control_OLS_Pool,
          title = "Table 4. OLS - Effects of full schoool days on test scores with controls",
          column.labels = c("Mathematics", "Language", "Pooled performance"),
          dep.var.labels = c("Test scores", "Test scores", "Test scores"),
          covariate.labels = c("Full-day school", "Public school", "Urban school", "School - SES 2", "School - SES 3", "School - SES 4", "Medium quality school", "High quality school", "Girl", "Student - SES 2", "Student - SES 3", "Student - SES 4", "Mother - full primary education", "Mother - some secondary education", "Mother - full secondary education", "Mother - some technical education", "Mother - full technical education", "Mother - some professional education", "Mother - full professional education", "Mother - masters or more", "Father - full primary education", "Father - some secondary education", "Father - full secondary education", "Father - some technical education", "Father - full technical education", "Father - some professional education", "Father - full professional education", "Father - masters or more"),
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis). Reference categories are Non-full day school, Private school, Rural School, School - SES 1, Low quality school, Boy, Student - SES 1, Mother - some primary education, Father - some primary education.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/Control_OLS.html")
```

## Model with quality-length interaction
```{r, warning=F}
Interaction_OLS_Mat <- lm(data = dfinal, Performance_Math ~ Full_School_Day:S_Quality + Sector_School + School_Area + School_SES  + Gender + Student_SES + Edu_Mother + Edu_Father)
Interaction_OLS_Lan <- lm(data = dfinal, Performance_Language ~ Full_School_Day:S_Quality + Sector_School + School_Area + School_SES  + Gender + Student_SES + Edu_Mother + Edu_Father)
Interaction_OLS_Pool <- lm(data = dfinal, Performance_MatLang ~ Full_School_Day:S_Quality + Sector_School + School_Area + School_SES  + Gender + Student_SES + Edu_Mother + Edu_Father)

stargazer(Interaction_OLS_Mat, Interaction_OLS_Lan, Interaction_OLS_Pool,
          title = "Table 5. OLS - Effects of full school days on test scores with controls and interaction of school day length and quality",
          
          column.labels = c("Mathematics", "Language", "Pooled performance"),
          dep.var.labels = c("Test scores", "Test scores", "Test scores"),
          covariate.labels = c("Public school", "Urban school", "School - SES 2", "School - SES 3", "School - SES 4", "Girl", "Student - SES 2", "Student - SES 3", "Student - SES 4", "Mother - full primary education", "Mother - some secondary education", "Mother - full secondary education", "Mother - some technical education", "Mother - full technical education", "Mother - some professional education", "Mother - full professional education", "Mother - masters or more", "Father - full primary education", "Father - some secondary education", "Father - full secondary education", "Father - some technical education", "Father - full technical education", "Father - some professional education", "Father - full professional education", "Father - masters or more", "Non-full day school - Low quality school", "Full day school - Low quality school", "Non-full day school - Medium quality school", "Full day school - Medium quality school", "Non-full day school - High quality school", "Full day school - High quality school"),
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Only the coefficients for the interaction are shown. Beta coefficients and standard errors (in parenthesis). Reference categories are Private school, Rural School, School - SES 1, Boy, Student - SES 1, Mother - some primary education, Father - some primary education.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/Interaction_OLS.html")
```

## Ordered probit models

## Ordered Probit - Na誰ve estimators
```{r}
Naive_OLS_Mat_PROBIT <- polr(data = dfinal, (as.factor(Performance_Math)) ~ Full_School_Day, Hess=TRUE)
Naive_OLS_Lan_PROBIT <- polr(data = dfinal, (as.factor(Performance_Language)) ~ Full_School_Day, Hess=TRUE)

stargazer(Naive_OLS_Mat_PROBIT, Naive_OLS_Lan_PROBIT,
          title = "Appendix 2a. Ordered probit - Effects of full school days on test scores (na誰ve estimators)",
          column.labels = c("Mathematics", "Language"),
          dep.var.labels = c("Test scores", "Test scores"),
          covariate.labels = "Full-day schooling",
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis). Reference category is Non full-day schooling.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/Naive_OLS_PROBIT.html")
```

## Ordered Probit with controls
```{r}
Control_OLS_Mat_PROBIT <- polr(data = dfinal, 
                               (as.factor(Performance_Math)) ~ Full_School_Day + Sector_School + School_Area + School_SES + S_Quality +
                                 Gender + Student_SES + Edu_Mother + Edu_Father, Hess=TRUE)

Control_OLS_Lan_PROBIT <- polr(data = dfinal, 
                               (as.factor(Performance_Language)) ~ Full_School_Day + Sector_School + School_Area + School_SES + S_Quality +
                                 Gender + Student_SES + Edu_Mother + Edu_Father, Hess=TRUE)

Control_OLS_Pool_PROBIT <- polr(data = dfinal, 
                                (as.factor(Performance_MatLang)) ~ Full_School_Day + Sector_School + School_Area + School_SES + S_Quality +
                                  Gender + Student_SES + Edu_Mother + Edu_Father, Hess=TRUE) 

stargazer(Control_OLS_Mat_PROBIT, Control_OLS_Lan_PROBIT, Control_OLS_Pool_PROBIT,
          title = "Appendix 2b. Ordered Probit - Effects of full school days on test scores with controls",
          column.labels = c("Mathematics", "Language", "Pooled performance"),
          dep.var.labels = c("Test scores", "Test scores", "Test scores"),
          covariate.labels = c("Full-day school", "Public school", "Urban school", "School - SES 2", "School - SES 3", "School - SES 4", "Medium quality school", "High quality school", "Girl", "Student - SES 2", "Student - SES 3", "Student - SES 4", "Mother - full primary education", "Mother - some secondary education", "Mother - full secondary education", "Mother - some technical education", "Mother - full technical education", "Mother - some professional education", "Mother - full professional education", "Mother - masters or more", "Father - full primary education", "Father - some secondary education", "Father - full secondary education", "Father - some technical education", "Father - full technical education", "Father - some professional education", "Father - full professional education", "Father - masters or more"),
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis). Reference categories are Non-full day school, Private school, Rural School, School - SES 1, Low quality school, Boy, Student - SES 1, Mother - some primary education, Father - some primary education.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/Control_OLS_PROBIT.html")
```


# Fixed Effects models

## FE without controls
```{r, warning=F}
FixedEM <- plm(Performance_Math ~ Full_School_Day, data = dfinal, index=c("ID11"), model = "within")
FixedEL <- plm(Performance_Language ~ Full_School_Day, data = dfinal, index=c("ID11"), model = "within")
Pooled_Fixed <- plm(Performance_MatLang ~ Full_School_Day, data = dfinal, index=c("ID11"), model = "within")

stargazer(FixedEM, FixedEL, Pooled_Fixed,
          title = "Table 7. FE - Effects of a full day of schooling on test scores (without controls)",
          column.labels = c("Mathematics", "Language", "Pooled"), 
          model.numbers = FALSE,
          dep.var.labels = c("Test scores", "Test scores", "Test scores"),
          covariate.labels = "Full-day schooling",
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis).  Reference category is Non full-day schooling.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/FE.html")
```

## FE controlling for wave
```{r, warning=F}
Wave_FixedEM <- plm(Performance_Math ~ Full_School_Day + Wave, data = dfinal, index=c("ID11", "Wave"), model = "within")
Wave_FixedEL <- plm(Performance_Language ~ Full_School_Day + Wave, data = dfinal, index=c("ID11", "Wave"), model = "within")
Wave_Pooled_Fixed <- plm(Performance_MatLang ~ Full_School_Day + Wave, data = dfinal, index=c("ID11", "Wave"), model = "within")

stargazer(Wave_FixedEM, Wave_FixedEL, Wave_Pooled_Fixed,
          title = "Table 8. FE - Effects of a full day of schooling on test scores (controlling for wave)",
          column.labels = c("Mathematics", "Language", "Pooled"), 
          model.numbers = FALSE,
          dep.var.labels = c("Test scores", "Test scores", "Test scores"),
          covariate.labels = c("Full-day school", "Second Wave"),
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis).  Reference categories are Non full-day school and First wave.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/Wave_FE.html")
```

# FE First difference 
```{r, warning=F}
FDFixedEM <- plm(Performance_Math ~ Full_School_Day, data = dfinal, index=c("ID11"), model = "fd")
FDFixedEL <- plm(Performance_Language ~ Full_School_Day, data = dfinal, index=c("ID11"), model = "fd")
FDPooled_Fixed <- plm(Performance_MatLang ~ Full_School_Day, data = dfinal, index=c("ID11"), model = "fd")

stargazer(FDFixedEM, FDFixedEL, FDPooled_Fixed,
          title = "Appendix 3. FE - Effects of a full day of schooling on test scores (first difference)",
          column.labels = c("Mathematics", "Language", "Pooled"), 
          model.numbers = FALSE,
          dep.var.labels = c("Test scores", "Test scores", "Test scores"),
          covariate.labels = "Full-day school",
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis).  Reference category is Non full-day school.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/FEFirstDifference.html")
```


# Fixed effects for sub-groups - effect heterogeneity

## School area
```{r, warning=F}
data_urban <- subset(dfinal, dfinal$School_Area == "1 - Urban")
data_rural <- subset(dfinal, dfinal$School_Area == "0 - Rural")

uFixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_urban, index=c("ID11"), model = "within")
uFixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_urban, index=c("ID11"), model = "within")

rFixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_rural, index=c("ID11"), model = "within")
rFixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_rural, index=c("ID11"), model = "within")

stargazer(uFixedEM, uFixedEL, rFixedEM, rFixedEL,
          title = "Appendix 4a. FE - Effects of a full day of schooling on test scores by school area",
          column.labels = c("Urban Schools - Mathematics", "Urban Schools - Language", "Rural Schools - Mathematics", "Rural Schools - Language"), 
          model.numbers = FALSE,
          dep.var.labels = c("Test scores", "Test scores", "Test scores", "Test scores"),
          covariate.labels = "Full-day school",
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis).  Reference category is Non full-day school.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/UrbanRuralFE.html")
```

## School sector
```{r, warning=F}
data_public <- subset(dfinal, dfinal$Sector_School == "1 - Public")
data_private <- subset(dfinal, dfinal$Sector_School == "0 - Private")

pubFixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_public, index=c("ID11"), model = "within")
pubFixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_public, index=c("ID11"), model = "within")

priFixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_private, index=c("ID11"), model = "within")
priFixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_private, index=c("ID11"), model = "within")

stargazer(pubFixedEM, pubFixedEL, priFixedEM, priFixedEL,
          title = "Appendix 4b. FE - Effects of a full day of schooling on test scores by school sector",
          column.labels = c("Public Schools - Mathematics", "Public Schools - Language", "Private Schools - Mathematics", "Private Schools - Language"), 
          model.numbers = FALSE,
          dep.var.labels = c("Test scores", "Test scores", "Test scores", "Test scores"),
          covariate.labels = "Full-day school",
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis).  Reference category is Non full-day school.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/PublicPrivateFE.html")
```

## School SES levels
```{r, warning=F}
data_SES1 <- subset(dfinal, dfinal$School_SES == 1)
data_SES2 <- subset(dfinal, dfinal$School_SES == 2)
data_SES3 <- subset(dfinal, dfinal$School_SES == 3)
data_SES4 <- subset(dfinal, dfinal$School_SES == 4)
# data_SES5 <- subset(dfinal, dfinal$School_SES == 5)

SES1FixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_SES1, index=c("ID11"), model = "within")
SES1FixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_SES1, index=c("ID11"), model = "within")

SES2FixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_SES2, index=c("ID11"), model = "within")
SES2FixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_SES2, index=c("ID11"), model = "within")

SES3FixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_SES3, index=c("ID11"), model = "within")
SES3FixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_SES3, index=c("ID11"), model = "within")

SES4FixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_SES4, index=c("ID11"), model = "within")
SES4FixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_SES4, index=c("ID11"), model = "within")

# There is not enough variation within schools with an SES of 5 to render any results for these models.
# SES5FixedEM <- plm(Performance_Math ~ Full_School_Day + Wave, data = data_SES5, index=c("ID11", "Wave"), model = "within")
# SES5FixedEL <- plm(Performance_Language ~ Full_School_Day + Wave, data = data_SES5, index=c("ID11", "Wave"), model = "within")


stargazer(SES1FixedEM, SES1FixedEL, SES2FixedEM, SES2FixedEL, SES3FixedEM, SES3FixedEL, SES4FixedEM, SES4FixedEL,
          title = "Appendix 4c. FE - Effects of a full day of schooling on test scores by school SES",
          column.labels = c("SES 1 - Mathematics", "SES 1 - Language", "SES 2 - Mathematics", "SES 2 - Language", "SES 3 - Mathematics", "SES 3 - Language", "SES 4 - Mathematics", "SES 4 - Language"), 
          model.numbers = FALSE,
          dep.var.labels = c("Test scores", "Test scores", "Test scores", "Test scores", "Test scores", "Test scores", "Test scores", "Test scores"),
          covariate.labels = "Full-day school",
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis).  Reference category is Non full-day school.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/SchoolSESFE.html")
```

## Individual SES levels
```{r, warning=F}
data_indivSES1 <- subset(dfinal, dfinal$Student_SES == 1)
data_indivSES2 <- subset(dfinal, dfinal$Student_SES == 2)
data_indivSES3 <- subset(dfinal, dfinal$Student_SES == 3)
data_indivSES4 <- subset(dfinal, dfinal$Student_SES == 4)

indivSES1FixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_indivSES1, index=c("ID11", "Wave"), model = "within")
indivSES1FixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_indivSES1, index=c("ID11", "Wave"), model = "within")

indivSES2FixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_indivSES2, index=c("ID11", "Wave"), model = "within")
indivSES2FixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_indivSES2, index=c("ID11", "Wave"), model = "within")

indivSES3FixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_indivSES3, index=c("ID11", "Wave"), model = "within")
indivSES3FixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_indivSES3, index=c("ID11", "Wave"), model = "within")

indivSES4FixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_indivSES4, index=c("ID11", "Wave"), model = "within")
indivSES4FixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_indivSES4, index=c("ID11", "Wave"), model = "within")



stargazer(indivSES1FixedEM, indivSES1FixedEL, indivSES2FixedEM, indivSES2FixedEL, indivSES3FixedEM, indivSES3FixedEL, indivSES4FixedEM, indivSES4FixedEL,
          title = "Appendix 4d. FE - Effects of a full day of schooling on test scores by individual SES",
          column.labels = c("SES 1 - Mathematics", "SES 1 - Language", "SES 2 - Mathematics", "SES 2 - Language", "SES 3 - Mathematics", "SES 3 - Language", "SES 4 - Mathematics", "SES 4 - Language"), 
          model.numbers = FALSE,
          dep.var.labels = c("Test scores", "Test scores", "Test scores", "Test scores", "Test scores", "Test scores", "Test scores", "Test scores"),
          covariate.labels = "Full-day school",
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis).  Reference category is Non full-day school.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/IndividualSEFE.html")
```

## School Quality
```{r, warning=F}
data_high <- subset(dfinal, dfinal$S_Quality == "High")
data_medium <- subset(dfinal, dfinal$S_Quality == "Medium")
data_low <- subset(dfinal, dfinal$S_Quality == "Low")

hFixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_high, index=c("ID11"), model = "within")
hFixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_high, index=c("ID11"), model = "within")

mFixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_medium, index=c("ID11"), model = "within")
mFixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_medium, index=c("ID11"), model = "within")

lFixedEM <- plm(Performance_Math ~ Full_School_Day, data = data_low, index=c("ID11"), model = "within")
lFixedEL <- plm(Performance_Language ~ Full_School_Day, data = data_low, index=c("ID11"), model = "within")

stargazer(hFixedEM, hFixedEL, mFixedEM, mFixedEL, lFixedEM, lFixedEL,
          title = "Appendix 4e. FE - Effects of a full day of schooling on test scores by school quality",
          column.labels = c("High Quality Schools - Mathematics", "High Quality Schools - Language", "Medium Quality Schools - Mathematics", "Medium Quality Schools - Language", "Low Quality Schools - Mathematics", "Low Quality Schools - Language"), 
          model.numbers = FALSE,
          dep.var.labels = c("Test scores", "Test scores", "Test scores", "Test scores", "Test scores", "Test scores"),
         covariate.labels = "Full-day school",
          type = "text",
          t.auto = F,
          p.auto = F,
          notes = "Beta coefficients and standard errors (in parenthesis).  Reference category is Non full-day school.",
          notes.align = "l",
          out = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/SchoolQualityFE.html")
```
