---
title: "Descriptives"
output:
  word_document: default
  html_document: default
---

```{r setup, include=T, warning=F}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stargazer)
library(ggplot2)
library(descr)
library(sjPlot)
library(readxl)
library(stringr)
library(plm)
library(table1)
library(RColorBrewer)
library(lattice)
options(scipen=999)
```

# Load data
```{r, include=T, warning=F}
load("/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Data/Final Data.Rda")
```

# Color palette
```{r, include=T, warning=F}
display.brewer.pal(n = 9, name = 'GnBu')
brewer.pal(n = 9, name = "GnBu")
```

# Sample size
```{r, warning=F}
# Persons and person/years
length(unique(dfinal$ID11)) # 32.581
length(dfinal$ID11) # 65.162
```

# Sample Statistics Dependent Variable
```{r, warning=F}
# Language
Table1a <- table(dfinal$Full_School_Day, dfinal$Performance_Language)
Table2a <- prop.table(Table1a, 1)
Table3a <- as.data.frame(Table2a)
Table3a$Percent <- Table3a$Freq*100

Fig1 <- ggplot(Table3a, 
       aes(fill=Var1, x=Var2, y=Percent)) +
  geom_bar(stat="identity", position=position_dodge(width=1)) + 
  ylab("%") + 
  xlab("Performance") +
  ggtitle("Figure 1. Performance in language by length of school day") + 
  labs(caption = "Own calculations with data from ICFES") + 
  scale_fill_manual(name = "Full school day", 
                    labels = c("Full day", "Not full day"), 
                    values = c("#A8DDB5", "#7BCCC4"))

ggsave(Fig1, file = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/Figure1.png")



# Mathematics
Table1b <- table(dfinal$Full_School_Day, dfinal$Performance_Math)
Table2b <- prop.table(Table1b, 1)
Table3b <- as.data.frame(Table2b)
Table3b$Percent <- Table3b$Freq*100

Fig2 <- ggplot(Table3b, 
       aes(fill=Var1, x=Var2, y=Percent)) +
  geom_bar(stat="identity", position=position_dodge(width=1)) + 
  ylab("%") + 
  xlab("Performance") +
  ggtitle("Figure 2. Performance in mathematics by length of school day") + 
  labs(caption = "Own calculations with data from ICFES") + 
  scale_fill_manual(name = "Full school day", 
                    labels = c("Full day", "Not full day"), 
                    values = c("#A8DDB5", "#7BCCC4"))
                    
ggsave(Fig2, file = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/Figure2.png")
```

# Distribution performance by subject and pooled
```{r, warning=F}
table1::label(dfinal$Performance_Language) <- "Performance Language"
table1::label(dfinal$Performance_Math) <- "Performance Mathematics"
table1::label(dfinal$Performance_MatLang) <- "Pooled Performance"
Table3 <- table1::table1(~ Performance_Language + Performance_Math + Performance_MatLang, data = dfinal, 
               caption = "Table 3. Sample statistics for performance by subject",
               footnote = "Own calculations with data from ICFES")

# Pooled
Table1c <- table(dfinal$Full_School_Day, dfinal$Performance_MatLang)
Table2c <- prop.table(Table1c, 1)
Table3c <- as.data.frame(Table2c)
Table3c$Percent <- Table3c$Freq*100

Fig3 <- ggplot(Table3c, 
       aes(fill=Var1, x=Var2, y=Percent)) +
  geom_bar(stat="identity", position=position_dodge(width=1)) + 
  ylab("%") + 
  xlab("Performance") +
  ggtitle("Fig. 3 Pooled performance by length of school day") + 
  labs(caption = "Own calculations with data from ICFES") + 
  scale_fill_manual(name = "Full school day", 
                    labels = c("Full day", "Not full day"), 
                    values = c("#A8DDB5", "#7BCCC4"))
                    
ggsave(Fig3, file = "/Users/isabeldebrigard/Google Drive/Hertie/GitHub_isabeldebrigard/Thesis/Thesis data/Figures/Figure3.png")
```

# Sample Statistics Independent Variables
```{r, warning=F}
table1::label(dfinal$Full_School_Day) <- "Full School day"
table1::label(dfinal$Gender) <- "Gender"
table1::label(dfinal$Sector_School) <- "School sector"
table1::label(dfinal$School_SES) <- "School SES"
table1::label(dfinal$School_Area) <- "School area"
table1::label(dfinal$Student_SES) <- "Student SES"
table1::label(dfinal$Edu_Mother) <- "Education of the Mother"
table1::label(dfinal$Edu_Father) <- "Education of the Father"
table1::label(dfinal$S_Quality) <- "School Quality"
Table1 <- table1::table1(~ Full_School_Day  + Sector_School + School_Area + School_SES + S_Quality + Gender + Student_SES + Edu_Mother + Edu_Father , data = dfinal, 
               caption = "Table 1. Sample statistics for main independent variables",
               footnote = "Own calculations with data from ICFES")

Table1
```

# Distribution of different school lengths
```{r, warning=F}
table1::label(dfinal$Length_School_Day) <- "Length of school day"
Table2 <- table1::table1(~ Length_School_Day, data = dfinal,
                         caption = "Table 2. Proportion of students by length of school day",
               footnote = "Own calculations with data from ICFES")

Table2
```

# Distribution: Length of the school day
```{r}
# Count 
lenght_school_day <- table(dfinal$Length_School_Day) 

# Percentages
round(((prop.table(table(dfinal$Length_School_Day))) * 100), 2)


barplot(lenght_school_day, main = "Length of the school day",
        col=c("#7BCCC4"))
```

# Distribution Gender
```{r, warning=F}
# Count
gender <- table(dfinal$Gender)
school_day_gender <- table(dfinal$Gender, dfinal$Length_School_Day)

# Percentages
round(((prop.table(table(dfinal$Gender))) * 100), 2)

round(((prop.table(table(dfinal$Gender, dfinal$Length_School_Day), 2)) * 100),2) # column percentages

barplot(school_day_gender, main = "Length of the school day by gender", 
        beside = T, 
        legend = row.names(school_day_gender),
                           col=c("#A8DDB5", "#7BCCC4"))
```

# Distribution School Sector
```{r, warning=F}
# Count
type <- table(dfinal$Sector_School)
Sector_School <- table(dfinal$Sector_School, dfinal$Length_School_Day)

# Percentages
round(((prop.table(table(dfinal$Sector_School))) * 100), 2)

round(((prop.table(table(dfinal$Sector_School, dfinal$Length_School_Day), 2)) * 100), 2) # column percentages

barplot(Sector_School, main = "Length of the school day by type of school", 
        beside = T, 
        legend = row.names(Sector_School),
                           col=c("#A8DDB5", "#7BCCC4"))
```

# Distribution School SES
```{r, warning=F}
# Count
schoolSES <- table(dfinal$School_SES)
schoolSES_school <- table(dfinal$School_SES, dfinal$Length_School_Day)

# Percentages
round(((prop.table(table(dfinal$School_SES))) * 100), 2)

round(((prop.table(table(dfinal$School_SES, dfinal$Length_School_Day), 2)) * 100), 2) # column percentages

barplot(schoolSES_school, main = "Length of the school day by School SES", 
        beside = T, 
        legend = row.names(schoolSES),
                           col=c("#CCEBC5", "#A8DDB5", "#7BCCC4", "#4EB3D3", "#2B8CBE"))
```

# Distribution School area
```{r, warning=F}
# Count
area <- table(dfinal$School_Area)
area_school <- table(dfinal$School_Area, dfinal$Length_School_Day)

# Percentages
round(((prop.table(table(dfinal$School_Area))) * 100), 2)

round(((prop.table(table(dfinal$School_Area, dfinal$Length_School_Day), 2)) * 100), 2) # column percentages


barplot(area_school, main = "Length of the school day by area of school", 
        beside = T, 
        legend = row.names(area_school),
                           col=c("#A8DDB5", "#7BCCC4"))
```

# Distribution Student SES
```{r, warning=F}
# Count
studentSES <- table(dfinal$Student_SES)
studentSES_school <- table(dfinal$Student_SES, dfinal$Length_School_Day)

# Percentages
round(((prop.table(table(dfinal$Student_SES))) * 100), 2)

round(((prop.table(table(dfinal$Student_SES, dfinal$Length_School_Day), 2)) * 100), 2) # column percentages

barplot(studentSES, main = "Student SES",
        col=c("#A8DDB5"))
barplot(studentSES_school, main = "Length of the school day by Student SES", 
        beside = T, 
        legend = row.names(studentSES),
                           col=c("#CCEBC5", "#A8DDB5", "#7BCCC4", "#4EB3D3", "#2B8CBE"))
```

# Distribution Education of the Father
```{r, warning=F}
# Count
eduF <- table(dfinal$Edu_Father)
eduF_school <- table(dfinal$Edu_Father, dfinal$Length_School_Day)

# Percentages
round(((prop.table(table(dfinal$Edu_Father))) * 100), 2)

round(((prop.table(table(dfinal$Edu_Father, dfinal$Length_School_Day), 2)) * 100), 2) # column percentages

barplot(eduF, main = "Father's Education level",
        col=c("#A8DDB5"))
barplot(eduF_school, main = "Length of the school day by Father's education", 
        beside = T, 
        legend = row.names(eduF),
                           col=c("#F7FCF0", "#E0F3DB", "#CCEBC5", "#A8DDB5", "#7BCCC4", "#4EB3D3", "#2B8CBE", "#0868AC", "#084081"))
```

# Distribution Education of the Mother
```{r, warning=F}
# Count
eduM <- table(dfinal$Edu_Mother)
eduM_school <- table(dfinal$Edu_Mother, dfinal$Length_School_Day)

# Percentages
round(((prop.table(table(dfinal$Edu_Mother))) * 100), 2)

round(((prop.table(table(dfinal$Edu_Mother, dfinal$Length_School_Day), 2)) * 100), 2) # column percentages

barplot(eduM, main = "Mother's Education level",
        col=c("#A8DDB5"))
barplot(eduM_school, main = "Length of the school day by Mother's education", 
        beside = T, 
        legend = row.names(eduM),
                           col=c("#F7FCF0", "#E0F3DB", "#CCEBC5", "#A8DDB5", "#7BCCC4", "#4EB3D3", "#2B8CBE", "#0868AC", "#084081"))
```

# Distribution School Quality
```{r, warning=F}
# Count
quality <- table(dfinal$S_Quality)
quality_school <- table(dfinal$S_Quality, dfinal$Length_School_Day)

# Percentages
round(((prop.table(table(dfinal$S_Quality))) * 100), 2)

round(((prop.table(table(dfinal$S_Quality, dfinal$Length_School_Day), 2)) * 100), 2) # column percentages
round(((prop.table(table(dfinal$S_Quality, dfinal$Length_School_Day), 1)) * 100),2) # row percentages

barplot(quality, main = "School Quality",
        col=c("#A8DDB5"))
barplot(quality_school, main = "Length of the school day by school quality", 
        beside = T, 
        legend = row.names(quality),
                           col=c("#F7FCF0", "#A8DDB5", "#0868AC"))
```

# Standard deviations and effect sizes
```{r, warning=F}
math = dfinal$Performance_Math
language = dfinal$Performance_Language
pooled = dfinal$Performance_MatLang

sd(math)
sd(language)
sd(pooled)

# OLS + controls
math_OLS_controls <- round(0.132/sd(math), digits = 2)  # medium
lang_OLS_controls <- round(0.062/sd(language), digits = 2) # medium


# OLS + interactions
math_OLS_interactions_low <- round(0.631/sd(math), digits = 2) # large
lang_OLS_int_eractions_low <- round(0.442/sd(language), digits = 2) # large
math_OLS_interactions_high <- round(0.214/sd(math), digits = 2) # large
lang_OLS_interactions_high <- round(0.099/sd(language), digits = 2) # medium

# FE
math_FE <- round(0.355/sd(math), digits = 2) # large
lang_FE <- round(0.059/sd(language), digits = 2) # medium

# (Kraft, 2020) "less than 0.05 is Small, 0.05 to less than 0.20 is Medium, and 0.20 or greater is Large"
```